---
title: "BSBM Benchmark Plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(purrr)
library(tibble)
library(jsonlite)
library(xml2)
```

## BSBM: total runtime per query, GengoDB vs. Fuseki Jena

Loads the runs stored in `<timestamp>-sf<scale factor>-<gengodb|fuseki>` under
`results/`, compute the *total* runtime as `aqet * executecount`. Results are 
faceted by scale factor for query-by-query comparison.

```{r bsbm-preload}
results_dir <- "results"

run_dirs <- list.dirs(results_dir, recursive = FALSE, full.names = TRUE)
run_dirs <- run_dirs[grepl("-(gengodb|fuseki)$", basename(run_dirs))]

parse_run <- function(dir) {
  meta <- fromJSON(file.path(dir, "meta.json"))
  xml <- read_xml(file.path(dir, "result.xml"))
  query_nodes <- xml_find_all(xml, ".//queries/query")
  tibble(
    engine = meta$backend,
    scale_factor = meta$product_count,
    query = as.integer(xml_attr(query_nodes, "nr")),
    execute_count = as.integer(xml_text(xml_find_first(query_nodes, "executecount"))),
    aqet = as.numeric(xml_text(xml_find_first(query_nodes, "aqet")))
  ) %>%
    mutate(total_runtime = aqet * execute_count)
}

bsbm_data <- map_dfr(run_dirs, parse_run) %>%
  filter(execute_count > 0, !query %in% c(6, 9, 12)) %>%
  mutate(
    engine = recode(engine, "gengodb" = "GengoDB", "fuseki" = "Fuseki Jena"),
    engine = factor(engine, levels = c("GengoDB", "Fuseki Jena")),
    scale_factor = factor(
      scale_factor,
      levels = sort(unique(scale_factor)),
      labels = paste0("SF ", sort(unique(scale_factor)))
    ),
    query = factor(query, levels = sort(unique(query)))
  ) %>%
  filter(scale_factor %in% c("SF 1000", "SF 2000", "SF 4000", "SF 5000"))
```

```{r bsbm-plot, fig.width=13, fig.height=10}
engine_colors <- c("GengoDB" = "#3C2673", "Fuseki Jena" = "#006e12")

# Mark tiny bars with a numeric label
bsbm_data <- bsbm_data %>%
  mutate(bar_label = if_else(total_runtime < 1, sprintf("%.3f", total_runtime), ""))

bsbm_plot <- ggplot(bsbm_data, aes(x = query, y = total_runtime, fill = engine)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = bar_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4, angle = 90, hjust = 0, size = 2.6, color = "#52514e"
  ) +
  facet_wrap(~scale_factor, scales = "free_y") +
  scale_fill_manual(values = engine_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.28))) +
  labs(
    title = "BSBM total query runtime: GengoDB vs. Fuseki Jena",
    subtitle = "Total execution time per query (aqet times execution count), by scale factor",
    x = "Query number",
    y = "Total runtime (s)",
    fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold")
  )
bsbm_plot
```

## Custom LingoDB backend: shared loading and plotting

Visualize individual module runs stored as `<timestamp>-sf<scale factor>-custom`
under `results/custom/`.

```{r custom-load}
custom_dir <- "results/custom"

custom_run_dirs <- list.dirs(custom_dir, recursive = FALSE, full.names = TRUE)
custom_run_dirs <- custom_run_dirs[grepl("-sf[0-9]+-custom$", basename(custom_run_dirs))]

read_custom_run <- function(dir) {
  read.csv(file.path(dir, "results.csv"), stringsAsFactors = FALSE) %>%
    as_tibble() %>%
    select(file, status, executionTime_mean, total_mean) %>%
    mutate(scale_factor = as.integer(sub(".*-sf([0-9]+)-custom$", "\\1", basename(dir))))
}

custom_runs <- map_dfr(custom_run_dirs, read_custom_run) %>%
  filter(status == "OK") %>%
  mutate(compile_time_mean = total_mean - executionTime_mean)

to_phases <- function(wide) {
  bind_rows(
    wide %>% mutate(phase = "Compile time", time_ms = compile_time_mean),
    wide %>% mutate(phase = "Execution time", time_ms = executionTime_mean)
  ) %>%
    mutate(phase = factor(phase, levels = c("Compile time", "Execution time")))
}
```

```{r custom-plot-helper}
phase_colors <- c("Compile time" = "#2a78d6", "Execution time" = "#eb6834")
group_colors <- c("#3C2673", "#9d8fd1")

plot_phase <- function(data, phase_name, x, title, subtitle, x_lab = NULL,
                       fill = NULL, facet = NULL, label_fmt = "%.3f") {
  data <- filter(data, phase == phase_name)
  dodge <- position_dodge(width = 0.8)
  label_layer <- function(...) {
    geom_text(
      aes(label = sprintf(label_fmt, time_ms), ...),
      vjust = -0.4, size = 2.8, color = "#52514e",
      position = if (is.null(fill)) "identity" else dodge
    )
  }

  p <- ggplot(data, aes(x = .data[[x]], y = time_ms))
  if (is.null(fill)) {
    p <- p +
      geom_col(width = 0.55, fill = phase_colors[[phase_name]]) +
      label_layer()
  } else {
    p <- p +
      geom_col(aes(fill = .data[[fill]]), position = dodge, width = 0.7) +
      label_layer(group = .data[[fill]]) +
      scale_fill_manual(values = group_colors)
  }
  if (!is.null(facet)) {
    p <- p + facet_wrap(vars(.data[[facet]]), scales = "free_x", nrow = 1)
  }

  p +
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    labs(
      title = title, subtitle = subtitle, x = x_lab,
      y = paste0(phase_name, " (ms)"), fill = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      strip.text = element_text(face = "bold")
    )
}
```

## Sub-operator modules: PageRank and Triangle Search

PageRank vs. PageRank with Graph SubOps and Triangle Search vs. Triangle
Search with Graph SubOps.

```{r custom-modules-load}
# file -> which pair it belongs to, and its short label within that pair
module_variants <- tribble(
  ~file, ~pair, ~variant,
  "pagerank.mlir", "PageRank", "Base",
  "pagerank_gsubop.mlir", "PageRank", "Graph SubOps",
  "triangle_search.mlir", "Triangle Search", "Base",
  "triangle_search_gsubop.mlir", "Triangle Search", "Graph SubOps"
)

module_data <- custom_runs %>%
  inner_join(module_variants, by = "file") %>%
  to_phases() %>%
  mutate(
    pair = factor(pair, levels = c("PageRank", "Triangle Search")),
    variant = factor(variant, levels = c("Base", "Graph SubOps"))
  )
```

```{r custom-modules-compile, fig.width=8, fig.height=4.5}
custom_modules_plot <- plot_phase(module_data, "Compile time",
  x = "variant", facet = "pair",
  title = "Sub-operator modules: compile time",
  subtitle = "Mean compile time per run (ms)", label_fmt = "%.1f ms"
)
custom_modules_plot
```

```{r custom-modules-execution, fig.width=8, fig.height=4.5}
custom_modules_plot <- plot_phase(module_data, "Execution time",
  x = "variant", facet = "pair",
  title = "Sub-operator modules: execution time",
  subtitle = "Mean execution time per run (ms)", label_fmt = "%.3f ms"
)
custom_modules_plot
```

## Bound variable join across scale factors

Triple Join (SubOp, "Hash Join") vs. Triple Join (Cheap, "Indexed Join"), each
run at SF1000-5000.

```{r triple-join-load}
triple_join_variants <- tribble(
  ~file, ~variant,
  "triple_join_subop.mlir", "Hash Join",
  "triple_join_cheap.mlir", "Index Join"
)

triple_join_data <- custom_runs %>%
  inner_join(triple_join_variants, by = "file") %>%
  to_phases() %>%
  mutate(
    variant = factor(variant, levels = triple_join_variants$variant),
    scale_factor = factor(paste0("SF ", scale_factor),
      levels = paste0("SF ", sort(unique(scale_factor)))
    )
  )
```

```{r triple-join-compile, fig.width=10, fig.height=4.5}
triple_hashjoin_plot <- plot_phase(triple_join_data, "Compile time",
  x = "scale_factor", fill = "variant",
  title = "Bound variable join: compile time",
  subtitle = "Mean compile time per run (ms), by scale factor",
  x_lab = "Scale factor", label_fmt = "%.1f"
)
triple_hashjoin_plot
```

```{r triple-join-execution, fig.width=10, fig.height=4.5}
triple_indexjoin_plot <- plot_phase(triple_join_data, "Execution time",
  x = "scale_factor", fill = "variant",
  title = "Bound variable join: execution time",
  subtitle = "Mean execution time per run (ms), by scale factor",
  x_lab = "Scale factor", label_fmt = "%.3f"
)
triple_indexjoin_plot
```

## BSBM queries on the custom backend

The BSBM query mix run at SF1000 and SF5000. Queries 6, 9 and
12 were not run against the GengoDB backend, since they are 
not yet supported by the implementation.

```{r bsbm-custom-load}
bsbm_custom_data <- custom_runs %>%
  filter(grepl("^[0-9]+\\.sparql$", file)) %>%
  mutate(query = as.integer(sub("\\.sparql$", "", file))) %>%
  to_phases() %>%
  mutate(
    query = factor(query, levels = sort(unique(query))),
    scale_factor = factor(paste0("SF ", scale_factor),
      levels = paste0("SF ", sort(unique(scale_factor)))
    )
  )
```

```{r bsbm-custom-compile, fig.width=11, fig.height=4.5}
plot_phase(filter(bsbm_custom_data, scale_factor == "SF 5000"), "Compile time",
  x = "query",
  title = "BSBM queries direct execution: compile time",
  subtitle = "Mean compile time per query (ms, SF5000)",
  x_lab = "Query number", label_fmt = "%.0f"
)
```

```{r bsbm-custom-execution, fig.width=11, fig.height=4.5}
plot_phase(bsbm_custom_data, "Execution time",
  x = "query", fill = "scale_factor",
  title = "BSBM queries direct execution: execution time",
  subtitle = "Mean execution time per query (ms), by scale factor",
  x_lab = "Query number", label_fmt = "%.2f"
)
```
